Equalizer 2.0\\Convergence of a Parallel Rendering Framework

### 2017-04-03 paperIEEE.tex
More benchmarks
git-svn-id: https://dev.ifi.uzh.ch/vmml/svn/Equalizer16@32 db7c848b-7755-4df2-b763-8796797afa52

%----------------------------------------------------------------------
\begin{figure*}[ht]\center
  \includegraphics[width=2\columnwidth]{images/teaser} \\
  (a) \hfil \hfil (b) \hfil \hfil (c)
  \vspace{-2mm}
  \caption{Example Equalizer application use cases: (a) 192 Megapixel CAVE at
    KAUST running RTNeuron, (b) Immersive HMD with external tracked and
    untracked views running RTT DeltaGen for virtual car usability studies (c)
    Cave2 running a molecular visualization build using Omegalib.}
  \label{FIG_teaser}
\end{figure*}

@@

The initial version of Equalizer implemented sort-first (2D), sort-last (DB),
stereo (EYE) and multilevel decompositions \cite{EMP:09}. In the following we
present new decomposition modes and motivate their use case. \fig{fig:compounds}
provides an overview of the new modes. We describe their configuration in
same \textit{compound} framework introduced in \cite{EMP:09}

@@

\begin{figure}[ht]\center
  \includegraphics[width=\columnwidth]{EqDocs/Developer/ProgrammingGuide/images/netObject}
  \caption{\label{fNetObject}Communication between two Collage Objects}
\end{figure}
\fig{fNetObject} illustrates this architecture.

@@

The most obvious one is compression. Recently, many new compression algorithms
have been developed which exploit modern CPU architectures and deliver
compression rates well above one gigabyte per second. Collage uses the Pression
library~\cite{pression}, which provides an unified interface for a number of
compression libraries, such as FastLZ~\cite{jesperfast}, Snappy~\cite{snappy}
and ZStandard~\cite{zstd}. It also contains a custom, virtually zero-cost RLE
compressor. Pression also parallelizes the compression and decompression using
data decomposition.

@@

\begin{figure}[ht]\center
  \includegraphics[width=.49\columnwidth]{images/spikes512}\hfil
  \includegraphics[width=.49\columnwidth]{images/spikes1280}
  \caption{\label{fVolumes}Volumes used for object distribution benchmarks:
    spike frequencies of a three million neuron electrical simulation at
    437x240x512 (51MB, left) and 1092x600x1280 (800MB, right) resolution}
\end{figure}

The spike frequency volumes aggregate the number of spikes which happened within
a given voxel over a given time. The absolute spike count is renormalized to an
unsigned byte value (0..255) during creation. Higher densities in the volume
represent higher spiking activity in the voxel. Consequently, the volume is more
sparsely populated and has a more even distribution of values at higher
resolutions.

@@

\begin{figure}[ht]\center
  \includegraphics[width=\columnwidth]{images/compressor}
  \caption{\label{fCompressor}Data Compression for Generic Binary Data}
\end{figure}

@@

\begin{figure*}[ht]\center
  \includegraphics[width=\textwidth]{images/compressorDetail}
  \caption{\label{fCompressorDetail}Data Compression for ply data (left, David
    statue 2mm) and raw volume data shown in \fig{fVolumes} (middle and right)}
\end{figure*}

@@

\subsubsection{Model Distribution and Update}
Distribute large ply and raw model to 1..16 nodes using TCP, IB and RSP

Distribute a video texture (per frame)

>>>>

%----------------------------------------------------------------------
\begin{figure*}[ht]\center
  \includegraphics[width=2\columnwidth]{images/teaser} \\
  (a) \hfil \hfil (b) \hfil \hfil (c)
  \vspace{-2mm}
  \caption{Example Equalizer application use cases: (a) 192 Megapixel CAVE at
    KAUST running RTNeuron, (b) Immersive HMD with external tracked and
    untracked views running RTT DeltaGen for virtual car usability studies (c)
    Cave2 running a molecular visualization build based on Omegalib.}
  \label{FIG_teaser}
\end{figure*}

@@

The initial version of Equalizer implemented sort-first (2D), sort-last (DB),
stereo (EYE) and multilevel decompositions \cite{EMP:09}. In the following we
present new decomposition modes and motivate their use case. \fig{fig:compounds}
provides an overview of the new modes. We describe their configuration in the
\textit{compound} framework introduced in \cite{EMP:09}.

@@

\begin{figure}[ht]\center
  \includegraphics[width=\columnwidth]{EqDocs/Developer/ProgrammingGuide/images/netObject}
  \caption{\label{fNetObject}Communication between two Collage objects}
\end{figure}
\fig{fNetObject} illustrates this architecture.

@@

The most obvious one is compression. Recently, many new compression algorithms
have been developed which exploit modern CPU architectures and deliver
compression rates well above one gigabyte per second. Collage uses the Pression
library~\cite{pression}, which provides an unified interface for a number of
compression libraries, such as FastLZ~\cite{jesperfast}, Snappy~\cite{snappy}
and ZStandard~\cite{zstd}. It also contains a custom, virtually zero-cost RLE
compressor. Pression parallelizes the compression and decompression using
data decomposition. Note that this compression is generic, and implemented
transparently for the application. Applications can still use data-specific
compression and feed the compressed data to Collage, typically disabling the
generic compression.

@@

\begin{figure}[ht]\center
  \includegraphics[height=.37\columnwidth]{images/spikes512}\hfil
  \includegraphics[height=.37\columnwidth]{images/beechnut}
  \caption{\label{fVolumes}Volumes used for object distribution benchmarks:
    spike frequencies of a three million neuron electrical simulation at
    437x240x512 unsigned byte (51MB, left) and an MRI scan of a beechnut at
    1024x1024x1546 unsigned short (3GB, right) resolution}
\end{figure}

The spike frequency volumes aggregate the number of spikes which happened within
a given voxel over a given time. The absolute spike count is renormalized to an
unsigned byte value (0..255) during creation. Higher densities in the volume
represent higher spiking activity in the voxel.

@@

\begin{figure}[ht]\center
  \includegraphics[width=\columnwidth]{images/compressor}
  \caption{\label{fCompressor}Data compression for generic binary data}
\end{figure}

@@

\begin{figure*}[ht]\center
  \includegraphics[width=\textwidth]{images/compressorDetail}
  \caption{\label{fCompressorDetail}Data compression for ply data (left, David
    statue 2mm) and raw volumes shown in \fig{fVolumes} (middle and right)}
\end{figure*}

@@

\subsubsection{Model Distribution and Update}

In this section we present results on how data distribution and synchronization
performs in real-world applications. We extracted the existing data distribution
code in a PLY and a volume renderer into a benchmark application to measure the
time to initially map all the objects on the render client nodes, and to perform
a commit-sync of the full data set after mapping has been established. All
figures observe a noticable measurement jitter due to other services running on
the cluster. The details of the benchmark algorithm can be found in the implementation~\cite{eqObjectBench}.

David 2mm

\begin{figure*}[ht]\center
  \includegraphics[width=\textwidth]{images/spikesdist}
  \caption{\label{fSpikesDist}Object mapping (left )and data synchronization
    times (middle, detail view right) for the spike volume data set in
    \fig{fVolumes}, right}
\end{figure*}

Distribute large ply and raw model to 1..16 nodes using TCP, IB and RSP
